A non-contextual hidden variable model for Quantum Mechanics

### 2016-07-07 finalPaper.tex
Fixed some typographical issues; QM quantum mechanics like and fixed some proposition etc. related things

\AtBeginDocument{\providecommand\thmref[1]{\ref{thm:#1}}}
\AtBeginDocument{\providecommand\eqref[1]{\ref{eq:#1}}}
\AtBeginDocument{\providecommand\figref[1]{\ref{fig:#1}}}
\AtBeginDocument{\providecommand\secref[1]{\ref{sec:#1}}}
\AtBeginDocument{\providecommand\ssecref[1]{\ref{ssec:#1}}}

\RS@ifundefined{subref}
  {\def\RSsubtxt{section~}\newref{sub}{name = \RSsubtxt}}
  {}
\RS@ifundefined{thmref}
  {\def\RSthmtxt{theorem~}\newref{thm}{name = \RSthmtxt}}
  {}
\RS@ifundefined{lemref}
  {\def\RSlemtxt{lemma~}\newref{lem}{name = \RSlemtxt}}
  {}

@@

\affiliation{Indian Institute of Science Education and Research (IISER), Mohali}
\begin{abstract}
We propose a non-contextual hidden variable model, consistent with all predictions of QM (for discrete spaces). We reconsider the no-go theorems on hidden variables and conclude that the notion of contextuality is not a necessary feature of quantum mechanics. Motivated by consistency requirements between Bohmian Mechanics and the said theorems, we provide an alternative view. This is accomplished by identifying a new classical property, we call `multiplicativity'. Advantages of this view are illustrated by considering the implications of non-multiplicativity to non-locality.%and its relation to contextuality.
\end{abstract}
\maketitle

@@

Background: Einstein's \cite{EinsteinEPR} work on incompleteness of quantum mechanics and the subsequent seminal work of Bell, assessing compatibility of hidden variable models with quantum mechanics has proven to be repeatedly useful. In Bell's case, for instance, it gave a precise meaning to the foundational notion of locality. Further, it has been of pragmatic utility in the areas of security, randomness certification etc. While Bell showed that local hidden variable models can not describe nature, the Kochen Specker theorem \cite{KochenSpecker} (and related work \cite{Gleason,BellOnHiddenVariables,Peres,Mermin}) ruled out an even larger class of hidden variable models. The notion of contextuality was identified which also, in recent times has been harnessed for computation and cryptography [TOOD: add references]. Not all hidden variable models were, however, proved to be incompatible. Bohm's work \cite{Bohm1,Bohm2} is one such (in)famous example. This approach has allowed for an interpretation of quantum mechanics with precise trajectories.

Previous work {[}the essentials I have already discussed{]}: We take a closer look at the aforesaid no-go theorems and find that contextuality is in fact not a necessary feature of quantum mechanics.
This has been shown earlier by taking disturbances due to measurements into consideration \cite{NoContextuality}. In this paper, we construct a significantly simpler toy model for illustrating not only the aforesaid but to also delineate a complete alternative to contextuality. The toy model is a specific case of what we call a c-ingle variable theory. Its predictions match with those of quantum mechanics for systems described by arbitrary (discrete) Hilbert spaces.

Triangle: <TODO: ask arvind sir, see \figref{block}>

@@

Work and results: In \secref{context} we establish a notation and definitions to facilitate further discussions. In \secref{nonMultiplicativity} we introduce the notion of multiplicativity and that of sequential multiplicativity. We review the `contextuality' theorem and prove some results about multiplicativity. We construct an explicit non-contextual model in \ssecref{cingle} and apply it to the very proof of `contextuality' to illustrate non-multiplicativity. This is supplemented by a discussion about Bohmian Mechanics in \ssecref{nonContBohm} where contextuality is removed by hand. The paper is concluded with discussion of implications of multiplicativity on locality, entanglement, superposition and apparent contextuality in \secref{implications}.

@@

\begin{notn} (a) $\psi\in\mathcal{H}$ would typically represent the quantum mechanical state of the system (assumed pure), residing in the Hilbert space $\mathcal{H}$, (b) $\hat{\mathcal{H}}$ is defined to mean $\mathcal{H}\otimes\mathcal{H}^{\dagger}$ (i.e. the set of Hermitian operators, observables, on the space $\mathcal{H}$), (c) $[\mathcal{H}]$ is defined to mean $(\mathcal{H},\,\mathbb{R}^{\otimes})$, which represents the state of the system including hidden variables, (d) $[\psi]\in[\mathcal{H}]$ will represent the state of the system, including hidden variables, (e) a prediction map is $m:\hat{\mathcal{H}},[\mathcal{H}]\to\mathbb{R}$, (f) a sequence map is $s:\hat{\mathcal{H}},[\mathcal{H}],\mathbb{R}\to[\mathcal{H}]$, (g) the set of all experimental setups is denoted by $\mathcal{E}$, (h) a setup map is $e:\hat{\mathcal{H}}\to\mathcal{E}$, (i) number of observables is represented by $N \in \mathbb{Z}^+$, (j) $f$ is an arbitrary map from $\hat{\mathcal{H}}^{\otimes N}\to \hat{\mathcal{H}}$ constructed using multiplication and addition of the observables, and multiplication with complex numbers\footnote{Strictly, the map $f$ would depend on the operators to ensure preservation of Hermiticity; thus a given $f$ will be defined only for a subset of $\hat {\mathcal H}^{\otimes N}$.}, (k) $\tilde{f}$ is a map from $\mathbb{R}^{\otimes N} \to \mathbb{R}$ constructed by replacing observables in $f$ with real numbers\footnote{$\tilde f$ constructed this way should be written as $\tilde f:\mathbb{R}^{\otimes N} \to \mathbb{C}$. In fact the values should be restricted to the spectrum of the associated observables.}.
\end{notn}

In addition to the notation, we use the following definition of non-contextuality.
\begin{defn}
A theory is non-contextual, if it provides a map $m:\hat{\mathcal{H}},[\mathcal{H}]\to\mathbb{R}$ to explain measurement outcomes. A theory which is not non-contextual is contextual.\end{defn}
\begin{rem}
A map is non-contextual if it is of the form $m:\hat{\mathcal{H}},[\mathcal{H}]\to\mathbb{R}$; prediction maps are non-contextual.
\end{rem}

In the literature broader definitions have been suggested which declare a larger set of theories as non-contextual. For our purposes however, this restricted definition will suffice. The term contextual is used to suggest that the value an operator takes might depend on which other compatible observable it is measured with. Since quantum mechanics (QM) yields deterministic results only in specific situations, it is imperative to analyse the meaning of contextuality by `completing' QM. Hidden variable theories do precisely this and we have at our disposal the heretic theory of Bohm. Circumventing the details of the theory, we note its salient feature; with each particle, one associates precisely defined $(q,p)$ and a wavefunction $\psi$. The construction is such that the particles are guided by the wavefunction. Consequently, the outcomes of all experiment are predictable, granted the initial conditions of the system and the apparatus are known (the `hidden variables'). The essence of Bohmian Mechanics relevant here can be captured by the following theorem.

\begin{thm}
$\exists$ a (Bohmian) map $m_{B}:\hat{\mathcal{H}},[\mathcal{H}],\mathcal{E}\to\mathbb{R}$,
and a sequence map $s$, s.t. if $m_{B}$ \& $s$ are assumed to describe
the outcomes of measurements \& the resultant state respectively,
then they are consistent with all predictions of QM.
\end{thm}
Observe that given a Stern-Gerlach setup and the initial conditions (position \& wavefunction of the particle), according to the theorem, Bohmian mechanics can predict the value of the spin (say $\hat{\sigma}_{z}$) of the particle. Assume that the state of the particle was $(\left|0\right\rangle +\left|1\right\rangle )/\sqrt{2}$ and that the measurement outcome was spin up. Interestingly, it can be shown \cite{Detlef} that if the direction of the magnetic field gradient is flipped while everything else is unchanged, the measurement outcome will now be spin down. This emphasises the fact that $m_{B}$ is not a prediction map. More importantly it entails that $m_{B}$ is contextual. It must be stressed, however, that the contextuality in this case is not a statement about the context in terms of compatible observables. Here no two observables are compatible (excluding identity).

@@

\begin{align*}
& m(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi]) = \\
& \tilde f(m(\hat{B}_{1},[\psi]),m(\hat{B}_{2},[\psi]),\dots m(\hat{B}_{N},[\psi])),
\end{align*}
where $\hat{B}_{i}\in\hat{\mathcal{H}}$ are arbitrary mutually commuting observables and $[\psi]\in[\mathcal{H}]$. A non-\emph{multiplicative} map is one that is not multiplicative.
\end{defn}
Note that if $m$ is taken to represent the measurement outcome (in
QM), then for states of the system which are simultaneous eigenkets
of $\hat{B}_{i}$s, $m$ must clearly be multiplicative. It is, however,
not obvious that this property must always hold. To motivate this,
consider the state of two spin half particles, $\left|11\right\rangle :=\left|1\right\rangle \otimes\left|1\right\rangle $
written in the computational basis, $\hat{B}_{1}=\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}$,
$\hat{B}_{2}=\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}$ and $\hat{C}=\hat{B}_{1}\hat{B}_{2}=-\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}$,
we have $m(\hat{C})=-1$ while $m(\hat{B}_{1})=\pm1$ and $m(\hat{B}_{2})=\pm1$,
independently, according to QM, with probability half. Here multiplicativity doesn't seem to hold. Antithetically it is clear that if one first measures
$\hat{B}_{1}$ and subsequently measures $\hat{B}_{2}$, then the
product of the results must be $-1$. This is consistent with measuring
$\hat{C}$. To capture this property we define \emph{sequential multiplicativity}
as follows.
\begin{defn}
A prediction map $m$ is \emph{sequentially multiplicative} for a
given sequence map $s$, iff 
\begin{align*}
&m(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi_{1}])=\\
&\tilde f(m(\hat{B}_{1},[\psi_{k_{1}}]),m(\hat{B}_{2},[\psi_{k_{2}}]),\dots,m(\hat{B}_{N},[\psi_{k_{N}}])),
\end{align*}
where $\mathbf{k}=(k_{1},k_{2},\dots,k_{N})\in\{(1,2,3\dots N),(2,1,3\dots N)$ + permutations, making $N!$ terms$\}$,
$\hat{B}_{i}\in\hat{\mathcal{H}}$ are arbitrary mutually commuting
observables, $[\psi_{i}]\in[\mathcal{H}]$ and $[\psi_{k+1}]\equiv s(\hat{B}_{k},[\psi_{k}],m(\hat{B}_{k},[\psi_{k}]))$,
$\forall\,[\psi_{i}]$.
\end{defn}
We now discuss the `proofs of contextuality' in the purview of the
aforesaid.
\begin{thm}
Let a map $m:\hat{\mathcal{H}}\to\mathbb{R}$, be s.t. (a) $m(\hat{\mathbb{I}})=1$,
(b) $m(f(\hat{A}_{1},\hat{A}_{2},\dots))=\tilde f(m(\hat{A}_{1}),m(\hat{A}_{2}),\dots)$,
where $\hat{A}_{i}$ are arbitrary Hermitian operators. If $m$ is assumed to describe the outcomes of measurements, then no $m$ exists which is consistent with all predictions of Quantum Mechanics.\label{thm:det} \end{thm}
Here $m$ maybe viewed as a specific class of prediction maps, that
implicitly depends on the state $[\psi]$. It follows then that the
maps ruled out by the theorem are non-contextual. The set of maps
ruled out, however, can be enlarged by imposing fewer restrictions
on it. This is achieved by the following theorem.

\begin{thm}
Let a map $m:\hat{\mathcal{H}}\to\mathbb{R}$, be s.t. (a) $m(\hat{\mathbb{I}})=1$,
(b) $m(f(\hat{B}_{1},\hat{B}_{2},\dots))=\tilde f(m(\hat{B}_{1}),m(\hat{B}_{2}),\dots)$,
for any arbitrary function $f$, where $\hat{B}_{i}$ are mutually
commuting Hermitian operators. If $m$ is assumed to describe the
outcomes of measurements, then no $m$ exists which is consistent
with all predictions of Quantum Mechanics.\label{thm:KS}\end{thm}

% The following proof is not only more general, but also independent
% of the initial state of the system.
\begin{proof}
Peres Mermin ($\left|\mathcal{H}\right|\ge4$):

Consider the following set of operators 
\[
\hat{A}_{ij}\doteq\left[\begin{array}{ccc}
\hat{\mathbb{I}}\otimes\hat{\sigma}_{x} & \hat{\sigma}_{x}\otimes\hat{\mathbb{I}} & \hat{\sigma}_{x}\otimes\hat{\sigma}_{x}\\
\hat{\sigma}_{y}\otimes\hat{\mathbb{I}} & \hat{\mathbb{I}}\otimes\hat{\sigma}_{y} & \hat{\sigma}_{y}\otimes\hat{\sigma}_{y}\\
\hat{\sigma}_{y}\otimes\hat{\sigma}_{x} & \hat{\sigma}_{x}\otimes\hat{\sigma}_{y} & \hat{\sigma}_{z}\otimes\hat{\sigma}_{z}
\end{array}\right]
\]
which have the property that all operators along a row (and along
a column) commute. It is trivial to see that this holds for the first
two rows and the first two columns. To see that this holds also for
the last row and column, note the anti-commutation relation, $\{\hat{\sigma}_{x},\hat{\sigma}_{y}\}=0$
and that $\sigma_{z}=i\sigma_{y}\sigma_{x}$, which one may check
explicitly. Another property is that the product of rows (columns)
yield $\hat{R}_{i}=\mathbb{I}$ and $\hat{C}_{j}=\mathbb{I}\,(j\neq3)$,
$\hat{C}_{3}=-\mathbb{I}$, ($\forall\,i,j$) where $\hat{R}_{i}\equiv\prod_{j}\hat{A}_{ij}$,
$\hat{C}_{j}\equiv\prod_{i}\hat{A}_{ij}$. This can be verified easily
by using the aforesaid relations and the fact that $\sigma^{2}=1$,
for every Pauli matrix. Using this property of $\hat{R}_{i}$ and
$\hat{C}_{j}$, it is easy to show that no map $m$ with the stated
properties can exist. Let us assume on the contrary that $m$ exists.
From property (b) of the map, to get $m(\hat{C}_{3})=-1$ (used property
(a)), we must have an odd number of $-1$ assignments in the third
column. In the remaining columns, the number of $-1$ assignments
must be even for each column. Thus, in the entire square, the number
of $-1$ assignments must be odd. Let us use the same reasoning, but
along the rows. Since each $m(\hat{R}_{i})=1$, we must have even
number of $-1$ assignments along each row. Thus, in the entire square,
the number of $-1$ assignments must be even. We have arrived at a
contradiction and therefore we conclude that our assumption that a
consistent $m$ exists, must be wrong.
\end{proof}

\begin{proof}
Kochen Specker ($\left|\mathcal{H}\right|\ge3$): The original proof
given by \cite{KochenSpecker} is rather involved. Here we will satisfy
ourselves with assuming its validity and applying our constructions
to the aforesaid proof(s).
\end{proof}

\begin{rem}
One could in principle assume $m$, to be s.t. (a) $m(\hat{\mathbb{I}})=1$,
(b) $m(\alpha\hat{B}_{i})=\alpha m(\hat{B}_{i})$, for $\alpha\in\mathbb{R}$,
(c) $m(\hat{B}_{i}^{2})=m(\hat{B}_{i})^{2}$, (d) $m(\hat{B}_{i}+\hat{B}_{j})=m(\hat{B}_{i})+m(\hat{B}_{j})$,
to deduce (d') $m(\hat{B}_{i}\hat{B}_{j})=m(\hat{B}_{i})m(\hat{B}_{j})$
and that $m(\hat{B}_{i})\in$ spectrum of $\hat{B}_{i}$. Effectively
then, condition (b) listed in the theorem is satisfied as a consequence.
Therefore, assuming (a)-(d) as listed here, rules out a larger class
of $m$.
\end{rem}
According to \thmref{KS}, non-contextual maps which are \emph{multiplicative}
must be incompatible with QM. It is apparent that non-contextual maps
which are non-\emph{multiplicative} might be consistent with QM. To
that end, in the following section we first show that this is indeed
the case by constructing an explicit model and supplement it by removing contextuality from Bohmian Mechanics.

>>>>

\AtBeginDocument{\providecommand\thmref[1]{\ref{thm:#1}}}
\AtBeginDocument{\providecommand\lemref[1]{\ref{lem:#1}}}
\AtBeginDocument{\providecommand\propref[1]{\ref{prop:#1}}}
\AtBeginDocument{\providecommand\eqref[1]{\ref{eq:#1}}}
\AtBeginDocument{\providecommand\figref[1]{\ref{fig:#1}}}
\AtBeginDocument{\providecommand\secref[1]{\ref{sec:#1}}}
\AtBeginDocument{\providecommand\ssecref[1]{\ref{ssec:#1}}}

\RS@ifundefined{subref}
  {\def\RSsubtxt{section~}\newref{sub}{name = \RSsubtxt}}
  {}
\RS@ifundefined{thmref}
  {\def\RSthmtxt{theorem~}\newref{thm}{name = \RSthmtxt}}
  {}
\RS@ifundefined{lemref}
  {\def\RSlemtxt{lemma~}\newref{lem}{name = \RSlemtxt}}
  {}
\RS@ifundefined{propref}
  {\def\RSproptxt{proposition~}\newref{prop}{name = \RSproptxt}}
  {}

@@

\affiliation{Indian Institute of Science Education and Research (IISER), Mohali}
\begin{abstract}
We propose a non-contextual hidden variable model, consistent with all predictions of quantum mechanics (QM). We reconsider the no-go theorems on hidden variables and conclude that the notion of contextuality is not a necessary feature of QM. Motivated by consistency requirements between bohmian mechanics and the said theorems, we provide an alternative view. This is accomplished by identifying a new classical property, we call `multiplicativity'. It turns out that if this condition is relaxed one is able to construct hidden variable models which are consistent with QM. Advantages of this view are illustrated by considering the implications of non-multiplicativity to non-locality. Our model is for finite dimensional quantum systems, however, generalization to the infinite dimensional case should be possible. %and its relation to contextuality.
\end{abstract}
\maketitle

@@

Einstein's \cite{EinsteinEPR} work on incompleteness of QM and the subsequent seminal work of Bell \cite{Bell1964}, assessing compatibility of hidden variable (HV) models with QM has proven to be repeatedly useful. In Bell's case, for instance, it gave a precise meaning to the foundational notion of locality. Further, it has been of pragmatic utility in the areas of security, randomness certification etc. \cite{Ekert,PironioRndmnssCrtfcn} While Bell showed that local HV models can not describe nature, the Kochen Specker theorem \cite{KochenSpecker} (and related work \cite{Gleason,BellOnHiddenVariables,Peres,Mermin}) ruled out an even larger class of HV models. Consequently the notion of contextuality was identified, methods to experimentally test it were proposed and evidence in its favour reported \cite{SimonContExpProp,HuangContExp,YangContExp,HasegawaContExp}. Contextuality also, in recent times has been harnessed for computation and cryptography \cite{HowardCntxCmptn,CabelloCntxScrt}. []Contextuality has also stirred further investigation of foundational aspects of QM: relation between contextuality and classicality has been explored \cite{PawelCntxClsscl}, contextuality has been quantified in terms of memory \cite{MatthiasCntxMmry} and bounded memory in conjunction with contextuality has been used to show completeness of QM \cite{CabelloMmryQM}.[] Not all HV models were, however, proved to be incompatible by the aforesaid theorem. Bohm's work \cite{Bohm1,Bohm2} is one such (in)famous example. This approach has allowed for an interpretation of QM with precise trajectories.

%Previous work {[}the essentials I have already discussed{]}: 
We take a closer look at the aforesaid no-go theorems and find that contextuality is in fact not a necessary feature of QM (see \figref{block}). This has been shown earlier by taking disturbances due to measurements into consideration \cite{NoContextuality,LaCourNoCntx}. In this paper, we construct a significantly simpler toy model for illustrating not only the aforesaid but to also delineate a complete alternative to contextuality. The toy model is a specific case of what we call a c-ingle variable theory. Its predictions match with those of QM for systems described by arbitrary finite dimensional Hilbert spaces.

%Triangle: <TODO: ask arvind sir, see \figref{block}>

@@

Organization of this paper is as follows. In \secref{context} we establish a notation and definitions to facilitate further discussions. In \secref{nonMultiplicativity} we introduce the notion of multiplicativity and that of sequential multiplicativity. We review the `contextuality' theorem and prove some results about multiplicativity. We construct an explicit non-contextual model in \ssecref{cingle} and apply it to the very proof of `contextuality' to illustrate non-multiplicativity. This is supplemented by a discussion about Bohmian Mechanics in \ssecref{nonContBohm} where contextuality is removed by hand. The paper is concluded with discussion of implications of multiplicativity on locality, entanglement, superposition and apparent contextuality in \secref{implications}.

@@

\begin{notn} (a) $\psi\in\mathcal{H}$ would typically represent the quantum mechanical state of the system (assumed pure), residing in the Hilbert space $\mathcal{H}$, (b) $\hat{\mathcal{H}}$ is defined to mean $\mathcal{H}\otimes\mathcal{H}^{\dagger}$ (i.e. the set of Hermitian operators, observables, on the space $\mathcal{H}$), (c) $[\mathcal{H}]$ is defined to mean $(\mathcal{H},\,\mathbb{R}^{\otimes})$, which represents the state of the system including HVs, (d) $[\psi]\in[\mathcal{H}]$ will represent the state of the system, including HVs, (e) a prediction map is $m:\hat{\mathcal{H}},[\mathcal{H}]\to\mathbb{R}$, (f) a sequence map is $s:\hat{\mathcal{H}},[\mathcal{H}],\mathbb{R}\to[\mathcal{H}]$, (g) $f$ is an arbitrary map from $\{ \hat{\mathcal{H}},\hat{\mathcal{H}},\dots\hat{\mathcal{H}} \} \to \hat{\mathcal{H}}$ constructed using multiplication and addition of the observables, and multiplication with complex numbers\footnote{Strictly, the map $f$ would depend on the operators to ensure preservation of Hermiticity; thus a given $f$ will be defined only for a subset of $\{ \hat{\mathcal{H}},\hat{\mathcal{H}},\dots\hat{\mathcal{H}} \}$.}, (h) $\tilde{f}$ is a map constructed by replacing observables in $f$ with real numbers. 
%\footnote{$\tilde f$ constructed this way should be written as $\tilde f:\mathbb{R}^{\otimes N} \to \mathbb{C}$. In fact the values should be restricted to the spectrum of the associated observables.}.
\end{notn}

In addition to the notation, we use the following definition of non-contextuality.
\begin{defn}
A theory is non-contextual, if it provides a map $m: \{ \hat{\mathcal{H}},[\mathcal{H}] \} \to\mathbb{R}$ to explain measurement outcomes. A theory which is not non-contextual is contextual.\end{defn}
\begin{rem}
The notion of non-contextuality may be extended to maps. A map is non-contextual if it is of the form $m: \{ \hat{\mathcal{H}},[\mathcal{H}] \} \to\mathbb{R}$; prediction maps are non-contextual.
\end{rem}

In the literature broader definitions have been suggested which declare a larger set of theories as non-contextual. For our purposes however, this restricted definition will suffice. The term contextual is used to suggest that the value an operator takes might depend on which other compatible observable it is measured with.

@@

\begin{align*}
& m(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi]) = \\
& \tilde f(m(\hat{B}_{1},[\psi]),m(\hat{B}_{2},[\psi]),\dots m(\hat{B}_{N},[\psi])),
\end{align*}
where $\hat{B}_{i}\in\hat{\mathcal{H}}$ are arbitrary mutually commuting observables and $[\psi]\in[\mathcal{H}]$. A non-\emph{multiplicative} map is one that is not multiplicative.
\end{defn}
Note that if $m$ is taken to represent the measurement outcome (in
QM), then for states of the system which are simultaneous eigenkets
of $\hat{B}_{i}$s, $m$ must clearly be multiplicative. It is, however,
not obvious that this property must always hold. To motivate this,
consider the state of two spin half particles, $\left|11\right\rangle :=\left|1\right\rangle \otimes\left|1\right\rangle $
written in the computational basis, $\hat{B}_{1}=\hat{\sigma}_{x}\otimes\hat{\sigma}_{x}$,
$\hat{B}_{2}=\hat{\sigma}_{y}\otimes\hat{\sigma}_{y}$ and $\hat{C}=\hat{B}_{1}\hat{B}_{2}=-\hat{\sigma}_{z}\otimes\hat{\sigma}_{z}$.
We must have $m(\hat{C})=-1$ while $m(\hat{B}_{1})=\pm1$ and $m(\hat{B}_{2})=\pm1$,
independently, according to QM, with probability half. Here multiplicativity doesn't seem to hold. Antithetically it is clear that if one first measures
$\hat{B}_{1}$ and subsequently measures $\hat{B}_{2}$, then the
product of the results must be $-1$. This is consistent with measuring
$\hat{C}$. To capture this property we define \emph{sequential multiplicativity}
as follows.
\begin{defn}
A prediction map $m$ is \emph{sequentially multiplicative} for a
given sequence map $s$, iff 
\begin{align*}
&m(f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{N}),[\psi_{1}])=\\
&\tilde f(m(\hat{B}_{1},[\psi_{k_{1}}]),m(\hat{B}_{2},[\psi_{k_{2}}]),\dots,m(\hat{B}_{N},[\psi_{k_{N}}])),
\end{align*}
where $\mathbf{k}=(k_{1},k_{2},\dots,k_{N})\in\{(1,2,3\dots N),(2,1,3\dots N)$ + permutations, making $N!$ terms$\}$,
$\hat{B}_{i}\in\hat{\mathcal{H}}$ are arbitrary mutually commuting
observables, $[\psi_{i}]\in[\mathcal{H}]$ and $[\psi_{k+1}]\equiv s(\hat{B}_{k},[\psi_{k}],m(\hat{B}_{k},[\psi_{k}]))$,
$\forall\,[\psi_{i}]$.
\end{defn}
We now discuss the `proofs of contextuality' in the purview of the
aforesaid.
\begin{thm}
Let a map $m:\hat{\mathcal{H}}\to\mathbb{R}$, be s.t. (a) $m(\hat{\mathbb{I}})=1$,
(b) $m(f(\hat{A}_{1},\hat{A}_{2},\dots))=\tilde f(m(\hat{A}_{1}),m(\hat{A}_{2}),\dots)$,
where $\hat{A}_{i}$ are arbitrary Hermitian operators. If $m$ is assumed to describe the outcomes of measurements, then no $m$ exists which is consistent with all predictions of QM\label{thm:det} (see \ref{proofs} for proofs). \end{thm}
Here $m$ maybe viewed as a specific class of prediction maps, that
implicitly depends on the state $[\psi]$. It follows then that the
maps ruled out by the theorem are non-contextual. The set of maps
ruled out, however, can be enlarged by imposing fewer restrictions
on it. This is achieved by the following theorem.

\begin{thm}
Let a map $m:\hat{\mathcal{H}}\to\mathbb{R}$, be s.t. (a) $m(\hat{\mathbb{I}})=1$,
(b) $m(f(\hat{B}_{1},\hat{B}_{2},\dots))=\tilde f(m(\hat{B}_{1}),m(\hat{B}_{2}),\dots)$,
for any arbitrary function $f$, where $\hat{B}_{i}$ are mutually
commuting Hermitian operators. If $m$ is assumed to describe the
outcomes of measurements, then no $m$ exists which is consistent
with all predictions of QM (see \ref{proofs} for more proofs).\label{thm:KS}\end{thm}

% The following proof is not only more general, but also independent
% of the initial state of the system.
\begin{proof}
Peres Mermin ($\left|\mathcal{H}\right|\ge4$) \cite{Peres,Mermin}:

Consider the following set of operators 
\[
\hat{A}_{ij}\doteq\left[\begin{array}{ccc}
\hat{\mathbb{I}}\otimes\hat{\sigma}_{x} & \hat{\sigma}_{x}\otimes\hat{\mathbb{I}} & \hat{\sigma}_{x}\otimes\hat{\sigma}_{x}\\
\hat{\sigma}_{y}\otimes\hat{\mathbb{I}} & \hat{\mathbb{I}}\otimes\hat{\sigma}_{y} & \hat{\sigma}_{y}\otimes\hat{\sigma}_{y}\\
\hat{\sigma}_{y}\otimes\hat{\sigma}_{x} & \hat{\sigma}_{x}\otimes\hat{\sigma}_{y} & \hat{\sigma}_{z}\otimes\hat{\sigma}_{z}
\end{array}\right]
\]
which have the property that all operators along a row (or column) commute and that the product of rows (or columns) yield $\hat{R}_{i}=\mathbb{I}$ and $\hat{C}_{j}=\mathbb{I}\,(j\neq3)$, $\hat{C}_{3}=-\mathbb{I}$, ($\forall\,i,j$) where $\hat{R}_{i}\equiv\prod_{j}\hat{A}_{ij}$, $\hat{C}_{j}\equiv\prod_{i}\hat{A}_{ij}$. Let us assume that $m$ exists. From property (b) of the map, to get $m(\hat{C}_{3})=-1$ (as required by property (a)), we must have an odd number of $-1$ assignments in the third column. In the remaining columns, the number of $-1$ assignments must be even (for each column). Thus, in the entire square, the number of $-1$ assignments must be odd. Let us use the same reasoning, but along the rows. Since each $m(\hat{R}_{i})=1$, we must have even number of $-1$ assignments along each row. Thus, in the entire square, the number of $-1$ assignments must be even. We have arrived at a contradiction and therefore we conclude that our assumption that a consistent $m$ exists, must be wrong.
\end{proof}

\begin{rem}
One could in principle assume $m$, to be s.t. (a) $m(\hat{\mathbb{I}})=1$,
(b) $m(\alpha\hat{B}_{i})=\alpha m(\hat{B}_{i})$, for $\alpha\in\mathbb{R}$,
(c) $m(\hat{B}_{i}^{2})=m(\hat{B}_{i})^{2}$, (d) $m(\hat{B}_{i}+\hat{B}_{j})=m(\hat{B}_{i})+m(\hat{B}_{j})$,
to deduce (d') $m(\hat{B}_{i}\hat{B}_{j})=m(\hat{B}_{i})m(\hat{B}_{j})$
and that $m(\hat{B}_{i})\in$ spectrum of $\hat{B}_{i}$. Effectively
then, condition (b) listed in the theorem is satisfied as a consequence.
Therefore, assuming (a)-(d) as listed here, rules out a larger class
of $m$. \cite{KochenSpecker}
\end{rem}
According to \thmref{KS}, non-contextual maps which are \emph{multiplicative}
must be incompatible with QM. It is apparent that non-contextual maps
which are non-\emph{multiplicative} might be consistent with QM. To
that end, in the following section we first show that this is indeed
the case by constructing an explicit model and supplement it by removing contextuality from Bohmian Mechanics.

Before proceeding we note, however, that QM can enforce \emph{sequential multiplicativity}. This result will be used to assess accuracy of any HV model we might propose.
\begin{prop*}
Let the system be in a state, s.t. measurement of $\hat{C}$ yields
repeatable results (same result each time). Then according to QM,
\emph{sequential multiplicativity} holds, where $\hat{C}\equiv f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})$,
and $\hat{B}_{i}$ are as defined. \end{prop*}
\begin{proof}
Assume without loss of generality that $\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}$
are mutually compatible (commuting) and complete set of operators.
If say for instance the set is not complete, then one can add the
missing operators and label them as aforesaid. It follows that $\exists$
$\left|\mathbf{b}=\left(b_{1}^{(l_{1})},b_{2}^{(l_{2})},\dots b_{n}^{(l_{n})}\right)\right\rangle $
s.t. $\hat{B}_{i}\left|\mathbf{b}\right\rangle =b_{i}^{(l_{i})}\left|\mathbf{b}\right\rangle $,
where $l_{i}$ indexes the eigenvalues corresponding to $\hat{B}_{i}$
and that $\sum_{\mathbf{b}}\left|\mathbf{b}\right\rangle \left\langle \mathbf{b}\right|=\hat{\mathbb{I}}$.
Let the state of the system be given by $\left|\psi\right\rangle $
and it must be s.t. $\hat{C}\left|\psi\right\rangle =c\left|\psi\right\rangle $,
by assumption. For the statement to follow, one need only show that
$\left|\psi\right\rangle $ must be made of only those $\left|\mathbf{b}\right\rangle $s,
which satisfy $c=\tilde f(b_{1}^{(l_{1})},b_{2}^{(l_{2})},\dots b_{n}^{(l_{n})})$.
This is the crucial step. Proving this is so is trivial. We start
with $\hat{C}\left|\psi\right\rangle =c\left|\psi\right\rangle $
and take its inner product with $\left\langle \mathbf{b}\right|$
to get 
\begin{eqnarray*}
\left\langle \mathbf{b}\right|\hat{C}\left|\psi\right\rangle  & = & c\left\langle \mathbf{b}|\psi\right\rangle ,\\
\left\langle \mathbf{b}\right|f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})\left|\psi\right\rangle  & = & c\left\langle \mathbf{b}|\psi\right\rangle ,\\
\tilde f(b_{1}^{(l_{1})},b_{2}^{(l_{2})},\dots b_{n}^{(l_{n})})\left\langle \mathbf{b}|\psi\right\rangle  & = & c\left\langle \mathbf{b}|\psi\right\rangle .
\end{eqnarray*}
Also, we have $\left|\psi\right\rangle =\sum_{\mathbf{b}}\left\langle \mathbf{b}|\psi\right\rangle \left|\mathbf{b}\right\rangle $, from completeness. If we consider $\left|\mathbf{b}\right\rangle $s for which $\left\langle \mathbf{b}|\psi\right\rangle \neq0$, then we can conclude that indeed $c=\tilde f(b_{1}^{(l_{1})},b_{2}^{(l_{2})},\dots b_{n}^{(l_{n})})$. However, when $\left\langle \mathbf{b}|\psi\right\rangle =0$, viz. $\left|\mathbf{b}\right\rangle $s that are orthogonal to $\left|\psi\right\rangle $, then nothing can be said. We can thus conclude that $\left|\psi\right\rangle $ is made only of those $\left|\mathbf{b}\right\rangle $s that satisfy the required relation. That completes the proof.
\end{proof}
It is worth noting that in the PM case, where $\hat{R}_{i}$ and $\hat{C}_{j}$ are just $\pm\hat{\mathbb{I}}$, it follows that all states are their eigenstates. Consequently, for these operators \emph{sequential multiplicativity} must always hold.
